Supplement to: Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-07 paper.tex
changes to list section

\begin{table}

  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \texttt{0}, \texttt{+1},\texttt{-1}\\
    \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{nth}\\
    \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
    \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}  
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\texttt{c}$) = $\texttt{(+ (chr->str c))}\\
      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s))}\\\hspace{2.55cm}\texttt{ (chr->str c))}\\
      $f_2($\texttt{s,c}$) = $\texttt{((}$f_0$\texttt{ c) (}$f_1$\texttt{ (1+ 0) c s))} \\
      \hspace{1cm}($f_2$: \emph{Extract substring delimited by character})
    %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\texttt{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
  \end{table}
%% \begin{figure}

@@

\subsection{List functions}
Our list function domain consists of tasks which are solved by functions
that take as input an integer or a list of integers, and have as output
either a Boolean, an integer, or a list of integers. Examples of these
functions are in Table~\ref{listexamples}. For each function, we create a
task $x$ by generating 15 input/output examples used for testing whether a
program produces the correct output. Supplying many examples reduces
ambiguity in the task's function, ensuring solutions achieve the desired
concept.

\begin{table}
\centering
\begin{tabular}{| l | l | l |}
  \hline
  \emph{name} & \emph{input} & \emph{output} \\
  \hline
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
  has-4 & [4\, 5\, 7\, 4] & \texttt{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \hline
\end{tabular}
\caption{Examples from the domain of list functions.}
\label{listexamples}
\end{table}

We supply \system with the DSL outlined in Table~\ref{listdsl}.

\begin{table}
\centering
\begin{tabular}{| l | l |}
  \hline
  \emph{name} & \emph{type} \\
  \hline
    empty & \texttt{[$\alpha$]} \\
    singleton & \texttt{$\alpha$ $\rightarrow$ [$\alpha$]} \\
    concat & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$]} \\
    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$} \\
    \hline
    true & \texttt{bool} \\
    not & \texttt{bool $\rightarrow$ bool} \\
    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    \hline
    0, $\ldots$, 9 & \texttt{int} \\
    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    negate & \texttt{int $\rightarrow$ int} \\
    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    is-prime & \texttt{int $\rightarrow$ bool} \\
    is-square & \texttt{int $\rightarrow$ bool} \\
    range & \texttt{int $\rightarrow$ [int]} \\
    sort & \texttt{[int] $\rightarrow$ [int]} \\
    \hline
    sum & \texttt{[int] $\rightarrow$ int} \\
    reverse & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$]} \\
    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
    index & \texttt{int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$} \\
    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
    slice & \texttt{int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
  \hline
\end{tabular}
\caption{DSL for the domain of list function.}
\label{listdsl}
\end{table}

\begin{figure}
  \includegraphics[width=\columnwidth]{figures/list.eps}
\end{figure}

We found that using a less sophisticated but equally-capable DSL made common
patterns, such as summation, unlikely and unlearnable in a small enumeration
bound.

>>>>

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \texttt{0}, \texttt{+1},\texttt{-1}\\
    \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{nth}\\
    \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
    \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}  
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\texttt{c}$) = $\texttt{(+ (chr->str c))}\\
      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s))}\\\hspace{2.55cm}\texttt{ (chr->str c))}\\
      $f_2($\texttt{s,c}$) = $\texttt{((}$f_0$\texttt{ c) (}$f_1$\texttt{ (1+ 0) c s))} \\
      \hspace{1cm}($f_2$: \emph{Extract substring delimited by character})
    %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\texttt{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
  \end{table}
%% \begin{figure}

@@

\subsection{List functions}
Our list function domain consists of hand-designed concepts from which we
sample tasks. All tasks are solved by functions that take as input an
integer or a list of integers, and have as output either a Boolean, an
integer, a list of Booleans, or a list of integers.  Examples of these
tasks are in Table~\ref{listexamples}.

\begin{table}
\centering
\begin{tabular}{| l | l | l |}
  \hline
  \emph{name} & \emph{input} & \emph{output} \\
  \hline
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
  has-head-in-tail & [4\, 5\, 7\, 4] & \texttt{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \hline
\end{tabular}
\caption{Examples of tasks from our list function domain. These are simpler
  than most other tasks in our dataset which are too complex to warrant
  description here.}
\label{listexamples}
\end{table}

We take approaches from two different initial DSLs, \emph{base} and
\emph{rich}, and assess their performance. The base DSL starts with only
low-level primitives such as \texttt{mapi}, \texttt{reducei}, and
\texttt{if}, whereas the rich DSL includes some common structure that can be
built from these such as \texttt{all}, \texttt{filter}, and \texttt{slice}.
Both are theoretically capable of solving all tasks supplied by our dataset.
See Supplement for more detail on these DSLs.

% TODO: table of list results, show some induced primitives

### 2018-02-07 supplement.tex
changes to list section

We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.\ref{programGenerativeModel}. Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
\begin{align*}
  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
\end{align}
where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.

@@

>>>>

We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.\ref{programGenerativeModel}. Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
\begin{align*}
  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
\end{align*}
where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.

@@

\vfill\null\columnbreak

\section{Initial domain-specific languages}
\subsection{List functions}
DSL for the list function experiment domain is shown below. The lower
section with seven primitives designates those given to the \emph{rich} DSL
variant and not the \emph{base} variant. The rich variant is not supplied
the \texttt{if} primitive, because its primary intention is to enable the
learning of the \texttt{filter}, \texttt{index}, and \texttt{slice}
concepts.

\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
    empty & [$\alpha$] \\
    singleton & $\alpha$ $\rightarrow$ [$\alpha$] \\
    concat & [$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
    mapi & (int $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) \\
    & \qquad\qquad$\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$] \\
    reducei & ($\beta$ $\rightarrow$ int $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$)\\
    & \qquad\qquad$\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$ \\
    \hline
    true & bool \\
    not & bool $\rightarrow$ bool \\
    and & bool $\rightarrow$ bool $\rightarrow$ bool \\
    or & bool $\rightarrow$ bool $\rightarrow$ bool \\
    if & bool $\rightarrow$ $\alpha$ $\rightarrow$ $\alpha$ $\rightarrow$ $\alpha$ \\
    \hline
    0 $\cdots$ 5 & int \\
    + & int $\rightarrow$ int $\rightarrow$ int \\
    * & int $\rightarrow$ int $\rightarrow$ int \\
    negate & int $\rightarrow$ int \\
    mod & int $\rightarrow$ int $\rightarrow$ int \\
    eq? & int $\rightarrow$ int $\rightarrow$ bool \\
    gt? & int $\rightarrow$ int $\rightarrow$ bool \\
    is-prime & int $\rightarrow$ bool \\
    is-square & int $\rightarrow$ bool \\
    range & int $\rightarrow$ [int] \\
    sort & [int] $\rightarrow$ [int] \\
    \hline
    sum & [int] $\rightarrow$ int \\
    reverse & [$\alpha$] $\rightarrow$ [$\alpha$] \\
    all & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool \\
    any & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool \\
    index & int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$ \\
    filter & ($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
    slice & int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$] \\
  \hline
\end{tabular}
\end{center}
